---
title: "Multinomial multilevel logistic regression"
author: "Anton Kalén"
date: today
format:
  html:
    self-contained: true
    code-fold: true
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
options(scipen=999)

# Load libraries

library(here)
library(haven)
library(dplyr)
library(brms)
# library(mclogit)
library(tidyr)
library(tidybayes)

#set priors
priors <- c(
  prior(normal(0,2.5), class = "b", dpar = "muATMT"),
  prior(normal(0,2.5), class = "b", dpar = "muMTAT")
)

# Load data
data <- read_sav(here("data", "Datasett Paper 2 Effekt logr til Andreas3 No change.sav"))

data_clean <- data %>% 
  mutate(
    across(c(kjonn, Etn, Tsent), as_factor),
    across(starts_with("Endring"), haven::as_factor),
    across(contains("4kat"), \(x) relevel(x, ref = "MT - MT")),
    across(starts_with("3kat"), \(x) relevel(x, ref = "No change")),
    across(ses_fr:ses_mr, \(x) x - 3),
    alder = scale(alder),
    Int_arm_2r = factor(Int_arm_2r, levels = c(2,1), labels = c("Kontroll", "Intervensjon"))
  ) %>% 
  drop_na(kjonn, alder, Etn, ses_fr, ses_mr, Int_arm_2r)


```

# Method

`alder` is standardised, meaning that the odds-ratio is the change in odds for one standard deviation difference in age.
Intercepts are therefore odds-ratios at mean age.

`ses_fr` and `ses_mr` is treated as continous and centered on the middle category (3 of 5), meaning that the odds ratio is the change for moving up or down one category. 
Intercepts are therefore odds-ratio for category 3 ("Fagskole").

We use a categorical multilevel regression to model the odds-ratio of pertaining to a group other than the control group. The `MT-MT` outcome is used as reference level.
Cases containing missing data are deleted listwise for each model, meaning that the number of cases will vary between models.
The results are presented as median log-odds ratio with 95% credibility intervals, given by the highest-density interval.
The interpretation of the CI is that there is a 95% probability that the odds-ratio falls in this range.
The models are fitted in STAN through the `brms` package using 4 chains, 2000 warm-up and 2000 sampling iterations.

# Endring_s7_3kat

```{r, include=FALSE}
data_Endring_s7_3kat <- data_clean %>% 
  drop_na(Endring_s7_3kat)

model_Endring_s7_3kat <- brm(
  Endring_s7_3kat ~ kjonn + alder + Etn + ses_fr + ses_mr + Int_arm_2r + Tsent + (1|Idsk),
  family = categorical(),
  data = data_Endring_s7_3kat,
  iter = 4000,
  chains = 4,
  cores = 4,
  prior = priors,
  control = list(adapt_delta = .9),
  backend = "cmdstanr",
  file = here("models", "model_Endring_s7_3kat.rds")
)
```

Fixed effects: 

```{r}
model_Endring_s7_3kat %>% 
  gather_draws(`b.*`, regex = TRUE) %>% 
  mutate(.value = exp(.value)) %>% 
  median_hdci() %>% 
  select(.variable:.upper) %>% 
  knitr::kable(digits = 2)
```

Random effects:

```{r}
model_Endring_s7_3kat %>% 
  gather_draws(`sd.*`, regex = TRUE) %>% 
  median_hdci() %>% 
  select(.variable:.upper) %>% 
  knitr::kable(digits = 2)
```

# Endring_s8_3kat

```{r, include=FALSE}
data_Endring_s8_3kat <- data_clean %>% 
  drop_na(Endring_s8_3kat)

model_Endring_s8_3kat <- brm(
  Endring_s8_3kat ~ 1 + kjonn + alder + Etn + ses_fr + ses_mr + Int_arm_2r + Tsent + (1|Idsk),
  family = categorical(),
  data = data_Endring_s8_3kat,
  iter = 4000,
  chains = 4,
  cores = 4,
  prior = priors,
  control = list(adapt_delta = .9),
  backend = "cmdstanr",
  file = here("models", "model_Endring_s8_3kat.rds")
)
```

Fixed effects: 

```{r}
model_Endring_s8_3kat %>% 
  gather_draws(`b.*`, regex = TRUE) %>% 
  mutate(.value = exp(.value)) %>% 
  median_hdci() %>% 
  select(.variable:.upper) %>% 
  knitr::kable(digits = 2)
```

Random effects:

```{r}
model_Endring_s8_3kat %>% 
  gather_draws(`sd.*`, regex = TRUE) %>% 
  median_hdci() %>% 
  select(.variable:.upper) %>% 
  knitr::kable(digits = 2)
```


# Endring_s9_3kat

```{r, include=FALSE}
data_Endring_s9_3kat <- data_clean %>% 
  drop_na(Endring_s9_3kat)

model_Endring_s9_3kat <- brm(
  Endring_s9_3kat ~ 1 + kjonn + alder + Etn + ses_fr + ses_mr + Int_arm_2r + Tsent + (1|Idsk),
  family = categorical(),
  data = data_Endring_s9_3kat,
  iter = 4000,
  prior = priors,
  control = list(adapt_delta = .9),
  chains = 4,
  cores = 4,
  backend = "cmdstanr",
  file = here("models", "model_Endring_s9_3kat.rds")
)
```


Fixed effects: 

```{r}
model_Endring_s9_3kat %>% 
  gather_draws(`b.*`, regex = TRUE) %>% 
  mutate(.value = exp(.value)) %>% 
  median_hdci() %>% 
  select(.variable:.upper) %>% 
  knitr::kable(digits = 2)
```

Random effects:

```{r}
model_Endring_s9_3kat %>% 
  gather_draws(`sd.*`, regex = TRUE) %>% 
  median_hdci() %>% 
  select(.variable:.upper) %>% 
  knitr::kable(digits = 2)
```


# Endring_s10_3kat

```{r, include=FALSE}
data_Endring_s10_3kat <- data_clean %>% 
  drop_na(Endring_s10_3kat)

model_Endring_s10_3kat <- brm(
  Endring_s10_3kat ~ 1 + kjonn + alder + Etn + ses_fr + ses_mr + Int_arm_2r + Tsent + (1|Idsk),
  family = categorical(),
  data = data_Endring_s10_3kat,
  iter = 4000,
  chains = 4,
  cores = 4,
  prior = priors,
  control = list(adapt_delta = .95),
  backend = "cmdstanr",
  file = here("models", "model_Endring_s10_3kat.rds")
)
```


Fixed effects: 

```{r}
model_Endring_s10_3kat %>% 
  gather_draws(`b.*`, regex = TRUE) %>% 
  mutate(.value = exp(.value)) %>% 
  median_hdci() %>% 
  select(.variable:.upper) %>% 
  knitr::kable(digits = 2)
```

Random effects:

```{r}
model_Endring_s10_3kat %>% 
  gather_draws(`sd.*`, regex = TRUE) %>% 
  median_hdci() %>% 
  select(.variable:.upper) %>% 
  knitr::kable(digits = 2)
```

# Multivariate model

```{r, include=FALSE}
data_Endring <- data_clean %>% 
  drop_na(starts_with("Endring"))

model_Endring <- brm(
  bf(
    mvbind(
      Endring_s7_3kat, 
      Endring_s8_3kat, 
      Endring_s9_3kat, 
      Endring_s10_3kat
    ) ~ 1 + kjonn + alder + Etn + ses_fr + ses_mr + Int_arm_2r + Tsent + (1|p|Idsk)
  ),
  family = categorical(),
  data = data_Endring,
  iter = 4000,
  chains = 4,
  cores = 4,
  prior = prior(normal(0,2.5), class = "b"),
  control = list(adapt_delta = .95),
  backend = "cmdstanr",
  file = here("models", "model_Endring.rds")
)

```

```{r}
model_Endring %>% 
  gather_draws(`b.*`, regex = TRUE) %>% 
  mutate(.value = exp(.value)) %>% 
  median_hdci() %>% 
  separate(.variable, into = c("b", "mu", "Endring", "var"), sep = "_") |> 
  arrange(Endring) |> 
  select(mu:.upper) %>% 
  knitr::kable(digits = 2)
```

Random effects:

```{r}
model_Endring %>% 
  gather_draws(`sd.*`, regex = TRUE) %>% 
  median_hdci() %>% 
  separate(.variable, into = c("sd", "ldsk", "empty", "mu", "Endring", "var"), sep = "_") |> 
  select(mu, Endring, .value:.upper) %>% 
  knitr::kable(digits = 2)
```

# Diagnostics

```{r}
tibble(
  models = c(
    "model_Endring_s7_3kat",
    "model_Endring_s8_3kat",
    "model_Endring_s9_3kat",
    "model_Endring_s10_3kat",
    "model_Endring"
  ),
  divergent = c(
    nuts_params(model_Endring_s7_3kat) |> filter(Parameter == "divergent__", Value == 1) |> nrow(),
    nuts_params(model_Endring_s8_3kat) |> filter(Parameter == "divergent__", Value == 1) |> nrow(),
    nuts_params(model_Endring_s9_3kat) |> filter(Parameter == "divergent__", Value == 1) |> nrow(),
    nuts_params(model_Endring_s10_3kat) |> filter(Parameter == "divergent__", Value == 1) |> nrow(),
    nuts_params(model_Endring) |> filter(Parameter == "divergent__", Value == 1) |> nrow()
  ),
  max_rhat = c(
    max(rhat(model_Endring_s7_3kat)),
    max(rhat(model_Endring_s8_3kat)),
    max(rhat(model_Endring_s9_3kat)),
    max(rhat(model_Endring_s10_3kat)),
    max(rhat(model_Endring))
  ),
 min_neff = c(
   min(neff_ratio(model_Endring_s7_3kat)),
   min(neff_ratio(model_Endring_s8_3kat)),
   min(neff_ratio(model_Endring_s9_3kat)),
   min(neff_ratio(model_Endring_s10_3kat)),
   min(neff_ratio(model_Endring))
 ) 
) |> 
  knitr::kable(digits = 4)


```

